---
title: "Multi-Precision-Computing-in-R"
output: rmarkdown::html_vignette
author: David Helmy
documentclass: article
vignette: >
  %\VignetteIndexEntry{Multi-Precision-Computing-in-R}
  %\VignetteEngine{rmarkdown::render}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
        collapse = TRUE,
        comment = "#>"
)
```

```{r setup}
library(MPCR)
```

# MPCR

MPCR is an advanced package designed to provide R users with a customized data structure.
This package harnesses the combined strength of C++ and R, empowering users with high-performance computing capabilities.
Specifically tailored for researchers and data scientists working with multi or mixed-precision arithmetic,
MPCR serves as an invaluable tool for achieving efficient and accurate computations.

##### MPCR provides R users with two primary, tailor-made data structures:
- Normal matrix/vector with different precision allocation (16-bit (half-precision), 32-bit (single-precision), and 64-bit (double precision)).
- The Tile-Matrix layout is constructed based on the standard MPCR matrix, enabling the creation of a matrix with multiple tiles, each having a distinct precision.
___

## Features
- Creation of matrix/vector with different precision allocation (16-bit (half-precision), 32-bit (single-precision), and 64-bit (double precision)).
- Support for all operators, basic utilities, binary operations, casters and converts, mathematical operations, and linear algebra.
- Tile-Matrix layout build-on normal MPCR matrix, offering the creation of a matrix with multiple tiles with a different precision for each one.
- Support for three main linear tile-algorithms `potrf, gemm, and trsm`.
- Support for converters to MPCR-Tile matrix

More details are available in the package [manual](../inst/doc/MPCR-manual.pdf)
___


## Example
```{r example-code-1}
# creating MPCRTile matrix and performing tile-potrf

# creating an R matrix of double precision values
R_matrix <- matrix(c(1.21, 0.18, 0.13, 0.41, 0.06, 0.23,
                     0.18, 0.64, 0.10, -0.16, 0.23, 0.07,
                     0.13, 0.10, 0.36, -0.10, 0.03, 0.18,
                     0.41, -0.16, -0.10, 1.05, -0.29, -0.08,
                     0.06, 0.23, 0.03, -0.29, 1.71, -0.10,
                     0.23, 0.07, 0.18, -0.08, -0.10, 0.36), 6, 6)

# creating a vector of strings, each string represents the precision of its corresponding tile.
# column major indexing is assumed
precision_metadata_vector <- c("single","double", "single", "single", "double", "double","single" , "single","double")

# MPCR Tile matrix initialization with size 6 x 6 and tile size 2 x 2
MPCRTile_mat <- new( MPCRTile , 6, 6, 2, 2, R_matrix, precision_metadata_vector)

# Perform out of place tile cholesky decomposition
chol_matrix <- chol(MPCRTile_mat,overwrite_input = FALSE)
print(chol_matrix)
```

```{r example-code-2}
# MPCR object creation
x <- new(MPCR, 50, "single")

# converting R object to MPCR object
y <- as.MPCR(1:24, precision = "single") # vector will be created

# converting R object to MPCR object
z <- as.MPCR(1:24, nrow = 4, ncol = 6, precision = "single") # Matrix will be created
```

More examples are available in [here](../tests/R-tests)
___


#### Note:
The speedup of MPCR over R is because MPCR is using MKL blas instead of Rblas, offering parallel computation on the data.
Normally you can use MKL backend with normal R objects, however, switching blas backends on R is quite complex and needs a lot of modification on the environment itself,
but in our case MPCR is using MKL without any modification to the environment itself, offering high speed computations with minimal efforts from the user side.
